---
title: "Site C Point Count Data: QA and Basic Exploration"
output:
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

This script is for import, tidy and quality assurance checks of the survey data.

To keep referential integrity, it seems best to join the three tables (stations, surveys and obs).

That way the visits can be properly renumbered (although that really only matters for the visit analysis)

# Setup

```{r message=FALSE, include=FALSE}
# Packages
library(tidyverse)
library(readxl)
library(lubridate)
library(skimr)
library(kableExtra)
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
options(scipen = 999)

```

# Read and Tidy Data
Use the import script.
```{r}
source("import_and_prep/import_point_count_data.R")
```


# Check for errors

Check for duplicates in stations\$StationID.

```{r}
stations %>% 
  count(StationID) %>% 
  filter(n>1)
```

Check for duplicates in surveys\$StationID & Visit.

Note: There are some duplicates in 2011 data. Need to fix (2020-06-29).

```{r}
surveys %>% 
  count(StationID, Visit) %>% 
  filter(n>1)
```

Check to see if there is a match for StationID between stations and surveys. This can be done using a full join and then filter.

```{r}
# Join the two tables
StationID_match <- stations %>% 
  full_join(surveys, by = "StationID")
# Filter to check for StationID in `surveys` but not in `stations`.
StationID_match %>% filter(is.na(Year))
# Filter to check for StationID in `stations` but not in `surveys`.
StationID_match %>% filter(is.na(Visit))
```

Now check for match for StationID between 'surveys' and 'obs'. Note that stations and surveys must be clean before this is useful.

For the second "test", surveys that had no observations will also be listed.

```{r paged.print=TRUE}
StationID_match <- surveys %>% 
  full_join(obs, by = c("StationID", "Visit"))
# Filter to check for `StationID $ Visit` in `obs` but not in `surveys`.
StationID_match %>% filter(is.na(Date))
# Filter to check for `StationID $ Visit` in `surveys` but not in `obs`.
# NB: This might be because of no birds observations rather than data entry error.
StationID_match %>% filter(is.na(SpCode))
```

Validate species codes in obs table.

NONE is used to indicate a survey with no observations.

```{r}
obs %>% 
  left_join(BCbirds, by = "SpCode") %>% 
  filter(is.na(`English Name`))
```

Check missing values (nulls).

```{r}
is.null(stations)
is.null(surveys)
is.null(obs)
```

# Create summary tables to see if any values out of range

Map station coordinates for anything out of range.

```{r}
stations %>% 
  ggplot(aes(x = Easting, y = Northing, colour = Year)) + 
  geom_point()
```

Summary by number of visits.

```{r message=FALSE}
visit_sum <- surveys %>% 
  group_by(StationID, Year = year(Date)) %>% 
  summarize(Visits = n()) %>% 
  mutate(Visits = as_factor(Visits)) %>% 
  group_by(Year, Visits) %>% 
  summarize("Number of Locations" = n())

# Make a summary table
visit_sum %>% 
  pivot_wider(id_cols = "Visits", 
              names_from = "Year", 
              values_from = "Number of Locations") %>% 
  replace(is.na(.), 0) %>% 
  kable()

# Plot for presentation
visit_sum %>% 
  ggplot(aes(x = as.character(Year), y = `Number of Locations`,
             fill = Visits)) +
  geom_bar(stat = "identity") +
#  scale_fill_brewer(palette="reds") +
  xlab("Year")
```

```{r paged.print=TRUE}
# Survey dates and times.
table(year(surveys$Date))
table(month(surveys$Date))
table(hour(surveys$Time))

# Observations counts
table(obs$Count)
# Distances
table(obs$`Distance Category`)
```

